Performance portability of an intermediate-complexity atmospheric research model in coarray Fortran

### 2017-10-05 response.txt

Attached please find a revised version of our extended abstract accepted for presentation at PAW17.

Reviewer 1
----------
1. Comment: "[2] Please cite the OpenSHMEM 1.3 specification (Available at: http://openshmem.org)"
   Response: The specification is now reference 10.

2. Comment: "Figure 1: The animation is probably not a good idea for the paper. Why not show a few frames instead?"
   Response: Figure 1 now depicts one representative frame. Space does not permit multiple frames.

3. Comment: "Figures 2-: The 'X' marks are indistinguishable when printed. Please use different symbols.
            Please also show error bars rather than plotting over the same point multiple times."
   Response: The symbols have been changed to better differentiate distinct runs.  We did not perform a sufficient
             number of runs to estimate error bars and the different data points are in fact distinct.  For example,
             the close correspondence between the MPI_put and OSH_put points in Figure 2 stems results from separate
             runs in each case and likely relates to commonalities in the underlying hardware or in the MPI and
             OpenSHMEM implementations.

   Comment: "What about the Intel Fortran compiler? Seems like this compiler should be used on KNL, and it should
            support coarrays?"
   Response: Section 1.2 now details our experiences with the Intel compiler, which produces executable programs
             that crash or hang beyond small core counts.

Review 2
--------

1. Comment (a): "The animation in Figure 1 depicts atmospheric processes over North 105 America as simulated by ICAR".
                Which animation? I would suggest take out animation (whatever that is) and put just a figure or snapshots.
   Response: Figure 1 now depicts one representative frame. Space does not permit multiple frames.

2. Comment (b): "Gets: an image that needs data proactively gets the data from memory managed by a providing image
                without the providing image’s active involvement,": what is an image?
                I know what it is, but it is not defined anywhere before this. A reader that knows nothing about CAF
                should at least be able to read what images are.
   Response: The first paragraph in Section 1 now gives a brief overview of the CAF programming model, including the meaning of "image."

3. Comment (c): "Figure 3 (top) depicts results across multiple compilers, machines, and architectures.
                In this case, only “puts” are tested, and only for the 2000x2000 domain.": why only puts? need to justify
                this more.
4. Response: The second paragraph of Section 3 now explains the reasons only "puts" are shown for MPI.

   Comment (d): in figure 3, lower right: title says runtime comparison, but there is no comparison.
   Response: The word "comparison" has now been deleted.

5. Comment (e): line 392 "This work as funded in..." typo.  "This work was funded in"
   Response: This typographical error has been corrected.

Review 3
--------
Comment: "1.2 spelling of "alteranative" needs to be alternative."
Response: The spelling has been corrected.

Comment: "2.2 The sentence 'We compiled at NCAR using of the GCC...." is poorly worded. Restructure the whole sentence
Response: The cited sentence has been resphrased for greater clarity.

### 2017-10-05 body.tex
Complete reviewer-suggested edits

\section{Introduction}
\subsection{Motivation and Background}
Since the Fortran 2008 standard was published in 2010~\cite{iso2010information}, a handful of published studies
have provided encouraging assessments of Fortran's intrinsic parallel programming model, \gls{caf}, in research
applications running at scale~\cite{preissl2011multithreaded,garain2015comparing,mozdzynski2015partitioned}.   One of the
most attractive features of \gls{caf} is the ability to express parallel algorithms once within the confines of a single
language standard without embedding compiler directives or referencing communication layers that are not part of the
language standard (e.g., no \gls{mpi} or OpenMP).  In theory, this should leave the choice of communication layers
as a link-time decision.  Ideally, an application developer writes a standard, Fortran-only application, compiles it with any
standard-conforming compiler, and possibly reserves the option to incorporate any one (or more) of several communication layers
at link-time.

@@

\gls{icar} simulates the motion of the atmosphere at kilometer length scales and produces flow patterns with a fidelity that is
attractive to the hydrology community, which studies surface water.  The animation in Figure~\ref{figure:icar} depicts atmospheric
processes over North America as simulated by \gls{icar}.

\begin{figure}
   \vspace{-18pt}
   % Animate frames at 7 fps created via 'ffmpeg -i qvmovie.mp4 -r 7 frame%d.pngr', add control buttons,
   % and automatically play when the page is viewed:
   \vbox{\hspace{-24pt}
   %\animategraphics[width=\columnwidth,controls,autoplay]{7}{figures/icar/frames/frame}{1}{30}
   \includegraphics[width=1.1\columnwidth]{figures/icar/frames/frame22.png}
   }
   \caption{An animation of atmospheric flow of water vapor (blues) and resulting precipitation (green to red) simulated by the \href{https://github.com/gutmann/icar}{\gls{icar}}. {\it For the animated version of this graphic, please access https://github.com/sourceryinstitute/coarray-icar-paw17/releases/download/1.0/main.pdf and use Adobe Acrobat Reader 7.0 or later.}
\label{figure:icar}}
\end{figure}

@@

\subsection{Objectives}
The goal of this work is to evaluate several alternatives for compiling, linking, and executing the
same \gls{mini-app} \gls{caf} source code using the following technologies:
\begin{itemize}
  \item \gls{mpi}~\cite{mpiforum2016mpi} and OpenSHMEM~\cite{openshmem2016} communication layers,
  \item \gls{gcc} and Cray compilers, and
  \item Multi- and many-core processors.
\end{itemize}
Additionally we show performance differences between two one-sided data access patterns:
\begin{enumerate}
  \item Gets: an image that needs data proactively gets the data from memory managed by a providing
        image without the providing image's active involvement, and
  \item Puts: an image that is able to provide data proactively puts the data into memory managed by
        a receiving image without the receiving image's active involvement.
\end{enumerate}
Fortran semantics necessitates that gets block.  Puts are non-blocking.

Our results will inform our future decisions around the choice of compiler, communication
layer, platform, and data access patterns.

@@

The test case implemented here uses a classical, idealized hill.
An idealized 1000 m high mountain is generated in the middle of the domain defined by a sine curve.
The air is initialized to near 100\% relative humidity, and a background wind of approximately 10 m s$^{-1}$ is specified.

At every time step, the outer edges of the local domain are processed first,
then passed to their neighbors while the interior of the domain is processed.
This is tested using both a \gls{caf} ``put'' operation, in which one-sided communication is initialized asynchronously to send data to a neighbor,
and a \gls{caf} ``get'' operation, in which one-sided blocking communication is used to retrieve data from a neighbor.

@@

We performed the experiments presented here on two supercomputers at \gls{nersc},
located at \gls{lbnl}, and one supercomputer at \gls{ncar}.
One \gls{lbnl} system is ``Edison,'' a Cray XC30 featuring 5586 nodes with two sockets of 12-core Intel Xeon Processor E5-2695 v2 (``Ivy Bridge''), running at \num{2.4}~\si{\giga\hertz}.
The second \gls{lbnl} system is ``Cori,'' a Cray XC40 which contains \num{12076} compute nodes, spanning two architectures. \num{2388} nodes each have two sockets of 16-core Intel Xeon Processor E5-2698 v3 (``Haswell'') operating at \num{2.3}~\si{\giga\hertz}; the other \num{9688} nodes are single-socket, 68-core Intel Xeon Phi Processor 7250 (``Knights Landing'') at \num{1.4}~\si{\giga\hertz}.
Both Edison and Cori feature the same ``Aries'' interconnect, with a dragonfly topology.
All measurements reported here which ran on Cori were executed on the Xeon Phi nodes, running in the ``quadrant`` NUMA configuration, with the MCDRAM configured as a transparent cache to the DDR4 memory.

The \gls{ncar} system is ``Cheyenne,'' an SGI ICE XA Cluster with \num{4032} compute nodes, with two sockets of 18-core Intel Xeon Processor E5-2697 v4 (``Broadwell''), running at \num{2.3}~\si{\giga\hertz}.
Cheyenne uses a Mellanox EDR infiniband interconnect with a partial 9D Enhanced Hypercube single-plane topology.
We compiled coarray-\gls{icar} on the \gls{nersc} systems using the Cray Fortran compiler version 8.6.0.  We compiled
at \gls{ncar} using the Fortran front end of \gls{gcc} version 6.3, which uses the
OpenCoarrays \gls{abi}~\cite{fanfarillo2014opencoarrays} for to support \gls{caf}.  We tested two OpenCoarrays parallel runtime libraries:
\begin{enumerate}
  \item The default \gls{mpi} library using the SGI MPT \gls{mpi},
  \item The openshmem branch using OpenSHMEM.
\end{enumerate}
We distributed \gls{caf} images as follows:
\begin{itemize}
  \item 68 images per Xeon Phi node on Cori,
  \item 24 images per Xeon node on Edison, and
  \item 36 images per node on Cheyenne.
\end{itemize}
corresponding to one image per physical core in each case.

\section{Discussion of Results}
Figure~\ref{fig1-2} shows results with two domain sizes (500 x 500 and 2000 x 2000 grid cells) running on Cheyenne.
These tests compare scaling with blocking ``get'' and non-blocking ``put'' operations, and compare scaling using \gls{mpi} or OpenSHMEM.
The differences between \gls{mpi} and OpenSHMEM are minor, with OpenSHMEM performing only slightly better.
However, the differences between the ``get'' and ``put'' operations are substantial.
For the 500x500 domain, ``puts'' scale well to 1800 cores, then scale poorly to \num{3600} cores;
however, ``gets'' only scale well to \num{504} cores, scale poorly to \num{1224}, and no further improvement comes from up to \num{3600} cores.

\gls{mpi} ``gets'' are not reported because a problem with the \gls{mpi} implementation makes them several orders of magnitude less efficient.
Similarly, \gls{mpi} ``put'' results are not reported beyond \num{1800} cores because a problem in the \gls{mpi} implementation caused memory
allocations to become a bottleneck at higher process counts.
For the 2000x2000 domain, the same pattern is observed but scaling of ``puts'' remains strong up to 3600 cores; whereas
scaling of ``gets'' continues, albeit poorly, to \num{3600} cores.

Figure~\ref{fig3-4} (top) depicts results across multiple compilers, machines, and architectures.
In this case, only ``puts'' are tested, and only for the 2000x2000 domain.
The GNU compiler was used on Cheyenne (Xeons). The Cray compiler was used on Edison (Xeons) and Cori (KNL).
To aid interpretation, the scaling results are reported as a fraction of the ideal scaling for each machine in the top left.
These results show that the Cray compiler+system scales better than the gfortran+SGI system,
with 75\% efficiency at >10k cores, while on Cheyenne, only 55\% of ideal was achieved.
This also shows that the KNL system scales well out to large core counts (60\% ideal with \num{19200} cores),
but that the total runtimes are significantly slower than the equivalent runtimes on Xeons (top right).
The KNL performance might be improved in the future by implementing OpenMP threaded parallelism within a node.

>>>>

\section{Introduction}
\subsection{Motivation and Background}
Since the Fortran 2008 standard was published in 2010~\cite{iso2010information}, a Fortran program executes as
a fixed number of instances that the standard refers to as ``images.''  Each image has its own state, including
data and input/output streams, corresponding to a \gls{spmd} programming style.  Images execute asynchronously
except at program launch, termination, and program-specified synchronization points.  Images may define
so-called coarray data structures to provide one-sided access to data by other images in a \gls{pgas}.

A few published studies have provided encouraging assessments of using Fortran's \gls{pgas} features
in research applications running at scale~\cite{preissl2011multithreaded,garain2015comparing,mozdzynski2015partitioned}.
One of the most attractive characteristics of these features is the ability to express parallel algorithms in a single
language without embedding compiler directives or referencing communication layers that are not part of the
language standard (e.g., no \gls{mpi} or OpenMP).  In theory, this should delay choice of communication layers
to link-time.  Ideally, an application developer writes a standard, Fortran-only application, compiles it into object code,
and then links to one or more communication layers.

@@

\gls{icar} simulates the motion of the atmosphere at kilometer length scales and produces flow patterns with a fidelity that is
attractive to the hydrology community for studying surface water.  Figure~\ref{figure:icar} depicts results from  \gls{icar}
simulations over North America.

\begin{figure}
   \vspace{-18pt}
   % Animate frames at 7 fps created via 'ffmpeg -i qvmovie.mp4 -r 7 frame%d.pngr', add control buttons,
   % and automatically play when the page is viewed:
   \vbox{\hspace{-24pt}
   %\animategraphics[width=\columnwidth,controls,autoplay]{7}{figures/icar/frames/frame}{1}{30}
   \includegraphics[width=1.1\columnwidth]{figures/icar/frames/frame22.png}
   }
   \vspace{-24pt}
   \caption{A visualization of the atmospheric distribution of water vapor (blues) and the resulting precipitation (green to red) simulated by \href{https://github.com/gutmann/icar}{\gls{icar}}.%{\it For the animated version of this graphic, please access https://github.com/sourceryinstitute/coarray-icar-paw17/releases/download/1.0/main.pdf and use Adobe Acrobat Reader 7.0 or later.}
\label{figure:icar}}
\end{figure}

@@

\subsection{Objectives}
This evaluates alternatives for compiling, linking, and executing one
\gls{mini-app} \gls{caf} source code using the following technologies:
\begin{itemize}
  \item \gls{mpi}~\cite{mpiforum2016mpi} and OpenSHMEM~\cite{openshmem2016} communication layers,
  \item \gls{gcc} and Cray compilers, and
  \item Multi- and many-core processors.
\end{itemize}
We also tested the Intel Fortran compiler versions 16 and 17. The resulting executable
program crashes on Cray platforms and hangs at launch on when run on more than 16 cores on the SGI platform.

We show performance for two one-sided data access patterns:
\begin{enumerate}
  \item Gets: an image retrieves data from memory managed by another
        image without the providing image's involvement, and
  \item Puts: an image stores data in memory managed by
        another image without the receiving image's involvement.
\end{enumerate}
Fortran semantics necessitates that gets block.  Puts are non-blocking.

%Our results will inform our future decisions around the choice of compiler, communication
%layer, platform, and data access patterns.

@@

The test case described here uses an idealized hill: a sine curve representing
a 1000 m high mountain in the middle of the domain.
The air is initially near 100\% relative humidity with a background wind of approximately 10 m s$^{-1}$.

At every time step, the outer edges of the local domain are processed first,
then passed to their neighbors while the interior of the domain is processed.
This is tested using both a \gls{caf} ``put'' operation, in which one-sided communication is initialized asynchronously to send data to a neighbor,
and a \gls{caf} ``get'' operation, in which one-sided blocking communication retrieves data from a neighbor.

@@

We performed the experiments presented here on two supercomputers at \gls{nersc},
located at \gls{lbnl}, and one supercomputer at \gls{ncar}. The \gls{lbnl} systems are
\begin{itemize}

\item Edison: a Cray XC30 featuring 5586 nodes with two sockets of 12-core Intel Xeon Processor E5-2695 v2 (``Ivy Bridge''), running at \num{2.4}~\si{\giga\hertz}.
\item Cori: a Cray XC40 with \num{12076} compute nodes, spanning two architectures; \num{2388} nodes each have two sockets of 16-core Intel Xeon Processor E5-2698 v3 (``Haswell'') operating at \num{2.3}~\si{\giga\hertz}; the other \num{9688} nodes are single-socket, 68-core Intel Xeon Phi Processor 7250 (``Knights Landing'') at \num{1.4}~\si{\giga\hertz}.
\end{itemize}
Edison and Cori have the same ``Aries'' interconnect with a dragonfly topology.
All Cori measurements on the Xeon Phi nodes ran in the ``quadrant'' NUMA configuration with the MCDRAM configured as a transparent cache to the DDR4 memory.

The \gls{ncar} system is ``Cheyenne,'' an SGI ICE XA Cluster with \num{4032} compute nodes, each containing two sockets of 18-core Intel Xeon Processor E5-2697 v4 (``Broadwell''), running at \num{2.3}~\si{\giga\hertz}.
Cheyenne uses a Mellanox EDR Infiniband interconnect with a partial 9D Enhanced Hypercube single-plane topology.
We compiled coarray-\gls{icar} on the \gls{nersc} systems using the Cray Fortran compiler version 8.6.0.  We compiled
at \gls{ncar} using the \gls{gcc} version 6.3 Fortran front end, which uses the
OpenCoarrays \gls{abi}~\cite{fanfarillo2014opencoarrays} to support \gls{caf}.  We tested two OpenCoarrays parallel runtime libraries:
\begin{enumerate}
  \item The default \gls{mpi} library using the SGI MPT \gls{mpi},
  \item The recently released OpenSHMEM library.
\end{enumerate}
We distributed \gls{caf} images as follows:
\begin{itemize}
  \item 68 images per Xeon Phi node on Cori,
  \item 24 images per Xeon node on Edison, and
  \item 36 images per node on Cheyenne,
\end{itemize}
corresponding to one image per physical core in each case.

\section{Discussion of Results}
Figure~\ref{fig1-2} shows results with two domain sizes ($500 \times 500 \times 20$ and $2000 \times 2000 \times 20$ grid cells) running on Cheyenne.
These tests compare scaling with blocking ``get'' and non-blocking ``put'' operations, and compare scaling using \gls{mpi} or OpenSHMEM.
The differences between \gls{mpi} and OpenSHMEM are minor, with OpenSHMEM performing only slightly better.
However, the differences between the ``get'' and ``put'' operations are substantial.
For the $500 \times 500 \times 20$ domain, ``puts'' scale well to 1800 cores, then scale poorly to \num{3600} cores;
however, ``gets'' only scale well to \num{504} cores, scale poorly to \num{1224}, and no further improvement comes from up to \num{3600} cores.

\gls{mpi} ``gets'' are not reported because a problem with the \gls{mpi} implementation makes them several orders of magnitude less efficient.
Similarly, \gls{mpi} ``put'' results are not reported beyond \num{1800} cores because a problem in the \gls{mpi} implementation caused memory
allocations to become a bottleneck at higher process counts.
The $2000 \times 2000 \times 20$ domain exhibits the same pattern but the scaling of ``puts'' remains strong up to 3600 cores; whereas
scaling of ``gets'' continues, albeit poorly, to \num{3600} cores.

Figure~\ref{fig3-4} (top) depicts results across multiple compilers, machines, and architectures.
In this case, only ``puts'' are tested, and only for the $2000 \times 2000\times 20$ domain.
The GNU compiler was used on Cheyenne (Xeons). The Cray compiler was used on Edison (Xeons) and Cori (KNL).
To aid interpretation, the scaling results are reported as a fraction of the ideal scaling for each machine in the top left.
These results show that the Cray compiler+system scales better than the gfortran+SGI system,
with 75\% efficiency at >10k cores, while on Cheyenne, only 55\% of ideal was achieved.
This also shows that the KNL system scales well out to large core counts (60\% ideal with \num{19200} cores),
but that the total runtimes are significantly slower than the equivalent runtimes on Xeons (top right).
The KNL performance might be improved in the future by implementing OpenMP threaded parallelism within a node.

### 2017-10-05 glossary/acronyms.tex
Add Sec 1.2 Objectives; edit 2.1 Physics &Numerics
Edit Section 2.2 Compilers, runtimes, and hardware
Complete reviewer-suggested edits

\newacronym{caf}{CAF}{coarray Fortran}
\newacronym{mini-app}{mini-app}{mini-application}
\newacronym{ncar}{NCAR}{National Center for Atmospheric Research}
\newacronym{wrf-hydro}{WRF-Hydro}{Weather Research and Forcecasting -- Hydrological Model}
\newacronym{spmd}{SPMD}{Single-Program Multiple-Data}
\newacronym{url}{URL}{Uniform Resource Locator}
\newacronym{tdd}{TDD}{test-driven development}
\newacronym{upc++}{UPC++}{Unified Parallel C++}
\newacronym{upc}{UPC}{Unified Parallel C}
\newacronym{lbnl}{LBNL}{Lawrence Berkeley National Laboratoory}
\newacronym{ide}{IDE}{integrated development environment}
\newacronym{html}{HTML}{Hypertext Markup Language}
\newacronym{icar}{ICAR}{The Intermediate Complexity Atmospheric Research}
\newacronym{exaflops}{exaFLOPS}{$10^{18}$ floating-point operations per second}
\newacronym{hpc}{HPC}{high-performance computing}
\newacronym{nasa}{NASA}{National Aeronautics and Space Administration}
\newacronym{nersc}{NERSC}{National Energy Research Scientific Computing Center}
\newacronym{doe}{DOE}{Department of Energy}
\newacronym{us}{US}{United States}
\newacronym{oop}{OOP}{Object-Oriented Programming}
\newacronym{ood}{OOD}{Object-Oriented Design}
\newacronym{uml}{UML}{Unified Modeling Language}
\newacronym{psblas}{PSBLAS}{Parallel Sparse Basic Linear Algebra Subroutines}
\newacronym{gpu}{GPU}{Graphics Processing Unit}
\newacronym{gpgpu}{GPGPU}{General-Purpose Graphics Processing Unit}
\newacronym{simt}{SIMT}{Single-Instruction, Multiple-Threads}
\newacronym{simd}{SIMD}{Single-Instruction, Multiple-Data}
\newacronym{2d}{2D}{Two-Dimensional}
\newacronym{ram}{RAM}{Random-Access Memory}
\newacronym{cpu}{CPU}{Central Processing Unit}
\newacronym{adsm}{ADSM}{Asymmetric Distributed Shared Memory}
\newacronym{loc}{LOC}{Lines Of Code}
\newacronym{csr}{CSR}{Compressed Storage by Rows}
\newacronym{coo}{COO}{COOrdinate}
\newacronym{csc}{CSC}{Compressed Sparse Columns}
\newacronym{ell}{ELL}{ELLpack}
\newacronym{jad}{JAD}{JAgged Diagonals}
\newacronym{cuda}{CUDA}{Compute Unified Device Architecture}
\newacronym{gflops}{GFLOPS}{Giga (billions) of Floating-point Operations Per Second, or GigaFLOPS}
\newacronym{pde}{PDE}{Partial Differential Equation}
\newacronym{spmv}{SpMV}{Sparse Matrix-Vector multiplication}
\newacronym{ecc}{ECC}{Error-Correcting Code}
\newacronym{petsc}{PETSc}{the Portable Extensible Toolkit for Scientific Computation}
\newacronym{flame}{FLAME}{the Formal Linear Algebra Methods Environment}
\newacronym{cfd}{CFD}{computational fluid dynamics}

>>>>

\newacronym{pgas}{PGAS}{partitioned global address space}
\newacronym{abi}{ABI}{application binary interface}
\newacronym{gcc}{GCC}{GNU Compiler Collection}
\newacronym{caf}{CAF}{coarray Fortran}
\newacronym{mini-app}{mini-app}{mini-application}
\newacronym{ncar}{NCAR}{the National Center for Atmospheric Research}
\newacronym{wrf-hydro}{WRF-Hydro}{Weather Research and Forcecasting -- Hydrological Model}
\newacronym{spmd}{SPMD}{single-program, multiple-data}
\newacronym{url}{URL}{Uniform Resource Locator}
\newacronym{tdd}{TDD}{test-driven development}
\newacronym{upc++}{UPC++}{Unified Parallel C++}
\newacronym{upc}{UPC}{Unified Parallel C}
\newacronym{lbnl}{LBNL}{Lawrence Berkeley National Laboratoory}
\newacronym{ide}{IDE}{integrated development environment}
\newacronym{html}{HTML}{Hypertext Markup Language}
\newacronym{icar}{ICAR}{The Intermediate Complexity Atmospheric Research}
\newacronym{exaflops}{exaFLOPS}{$10^{18}$ floating-point operations per second}
\newacronym{hpc}{HPC}{high-performance computing}
\newacronym{nasa}{NASA}{National Aeronautics and Space Administration}
\newacronym{nersc}{NERSC}{the National Energy Research Scientific Computing Center}
\newacronym{doe}{DOE}{Department of Energy}
\newacronym{us}{US}{United States}
\newacronym{oop}{OOP}{Object-Oriented Programming}
\newacronym{ood}{OOD}{Object-Oriented Design}
\newacronym{uml}{UML}{Unified Modeling Language}
\newacronym{psblas}{PSBLAS}{Parallel Sparse Basic Linear Algebra Subroutines}
\newacronym{gpu}{GPU}{Graphics Processing Unit}
\newacronym{gpgpu}{GPGPU}{General-Purpose Graphics Processing Unit}
\newacronym{simt}{SIMT}{Single-Instruction, Multiple-Threads}
\newacronym{simd}{SIMD}{Single-Instruction, Multiple-Data}
\newacronym{2d}{2D}{Two-Dimensional}
\newacronym{ram}{RAM}{Random-Access Memory}
\newacronym{cpu}{CPU}{Central Processing Unit}
\newacronym{adsm}{ADSM}{Asymmetric Distributed Shared Memory}
\newacronym{loc}{LOC}{Lines Of Code}
\newacronym{csr}{CSR}{Compressed Storage by Rows}
\newacronym{coo}{COO}{COOrdinate}
\newacronym{csc}{CSC}{Compressed Sparse Columns}
\newacronym{ell}{ELL}{ELLpack}
\newacronym{jad}{JAD}{JAgged Diagonals}
\newacronym{cuda}{CUDA}{Compute Unified Device Architecture}
\newacronym{gflops}{GFLOPS}{Giga (billions) of Floating-point Operations Per Second, or GigaFLOPS}
\newacronym{pde}{PDE}{Partial Differential Equation}
\newacronym{spmv}{SpMV}{Sparse Matrix-Vector multiplication}
\newacronym{ecc}{ECC}{Error-Correcting Code}
\newacronym{petsc}{PETSc}{the Portable Extensible Toolkit for Scientific Computation}
\newacronym{flame}{FLAME}{the Formal Linear Algebra Methods Environment}
\newacronym{cfd}{CFD}{computational fluid dynamics}
